---
title: "data1"
output:
  pdf_document: default
  word_document: default
  html_document: default
---

```{r}
# 샘플링 패키지

library(doBy)
library(sampling)
library(dplyr)
library(ggplot2)
library(MASS)
library(car)
library(outliers)
library(DMwR)
library(Rlof)
library(nnet)
library(gmodels)
library(caret)
library(e1071)
library(randomForest)
library(KoNLP)
library(hflights)
library(readr)


```

# 데이터분석 1일차

### Sampling

```{r}
sample(1:100, 5, replace = T)

```

parameter in sample() - prob : get sample with weighted probability.

### Using the same seed for random number

```{r}
set.seed(100)
sample(1:100, 5, replace = T)
```

Now, r always returns the same random number from the seed.

```{r}
#View(iris)
#edit(iris)
```
View is for viewing dataset, and edit is editing.

When you look into the iris dataset, you can find out the data raw are sorted with species. Shuffle is needed to avoid this situation. overfitting now occurs.


```{r}
mat <- cbind(c(1,2,3),c(1,2,4))
mat
nrow(mat)
ncol(mat)
length(mat)
```

- nrow : row count
- ncol : column count
- length : the number of each element of all rows and columns. given 3 * 4, length() will return 12.

```{r}
a <- 100
b <<- 3
```

<- for allocating local variable
<<- for allocating global variable

### Simple random sampling

```{r}
set.seed(5882)
ind1 = sample(1:a, a*0.7, )
ind1
```

### Stratified sampling

```{r}
train = iris[1:5,2]
train = iris[c(1,3,100),2]
train = iris[-c(1,3,100),2]
train = iris[-c(1,3,100),]
train = iris[ind1, ]
train
```

### how to get sample for train and test data?

just use -(minus) operator.

```{r}
train = iris[ind1,]
test = iris[-ind1,]
```
  
### stratified sampling (층화추출, 계통추출)

```{r}
#ind2 <- sampleBy ( ~ Species, data = iris, frac=c(0.1), syetematic = T)
#ind2
```

with systematic options, it fetches data with same gap among each samples.

### stratified sampling (층화추출)

```{r}
#be1 = a/5
#st1 = sample(1:be1, 1)
#rep()
#seq(st1, a, be1)
#seq(start, end, step)
```
rep : get same number repeatdly
seq : get numbers between given option

### 집락추출

중학교 추출시, 한학교를 통채로 뽑아와서 그놈을 다 보는것. 다른학교 꺼지고 지역중 하나만 중학교를 뽑아서 보는것.
집단을 하나 그대로 와서 하는걸 집락추출.

```{r}
ind3 = strata(iris, 'Species', c(7,5,3), 'srswor')
getdata(iris,ind3)


```


get 7 from the first specis, 5 from the second, 3 from the third. 'srswor' for replcement logic.


```{r}
hist(iris$Sepal.Length)
mean(iris$Sepal.Length)
median(iris$Sepal.Length)
summary(iris)
```

### table
```{r}
table(iris$Species)
```
if numeric data should be in the range class, cut() can be used.

```{r}
cut(iris$Sepal.Length, 5)
cut(iris$Sepal.Length, c(0,5, 10, 12, 20))
g1 =cut(iris$Sepal.Length, c(0,5, 10, 12, 20), right = F)
```

right is the options which tells you to include the last element of the array.

```{r}
table(g1)
```

outlier affects mean too much.

### 표본분포

```{r}
xxbar = c()
for(i in 1:100) {
  ss1 = sample(1:10000, 10, rep = T)
  xbar = mean(ss1)
  xxbar = c(xxbar, xbar)
}
hist(xxbar)
mean(xxbar)
mean(1:100)
                                                     ```






